From Freebase to Wikidata: The Great Migration

### 2015-10-13 paper.tex
Reworked Sections 5.1 and 5.2.

\usepackage[utf8]{inputenc}
\usepackage[hyphens]{url}
\usepackage[pdftex,urlcolor=black,colorlinks=true,linkcolor=black,citecolor=black]{hyperref}
\def\sectionautorefname{Section}
\def\subsectionautorefname{Subsection}
\usepackage{graphicx}
% todo macro
\usepackage{color}
\newcommand{\todo}[1]{\noindent\textcolor{red}{{\bf \{TODO}: #1{\bf \}}}}

@@

Here is an overview of the size of the last dump of Freebase:

\begin{itemize}
  \setlength\itemsep{0em}
  \item 48M~topics
  \item 2,997M triples
  \item 442M facts owned by Google\footnote{These ``facts'' are the Freebase triples
      after having filterout the data about ids (\texttt{/type/object/id}),
      types (\texttt{/type/object/type}), labels and descriptions (\texttt{/type/object/name},
      \texttt{/common/topic/description}...) and triples that should be filtered
      for legal reasons (see \autoref{sec:licensing}).}
  \item 68M~labels
\end{itemize}

To compare, Wikidata contains (as of August 2015):

@@

But why are these numbers so different?
Regarding the number of topics, Freebase has a~lot of topics about subjects
that do not match the Wikidata notability criteria%
\footnote{\url{https://www.wikidata.org/wiki/Wikidata:Notability}}.
For example, Freebase has 11.3M~musical recordings, 8.3M~music canonical versions,
2.6M~ISBNs, \emph{etc.}
We also cannot compare the number of facts and the number of statements directly,
even considering the lower number of topics.
Freebase encodes its knowledge in far many more facts than Wikidata needs statements.
Properties in Freebase often have a~reverse property used in order to be able
to traverse the Freebase graph easily in both directions.
For example the property \texttt{/people/person/place\_of\_birth} will have as reverse
the property \texttt{/location/location/people\_born\_here} that encodes
exactly the same semantic information.
Such reverses sometimes exist in Wikidata (like ``children'' for ``father'' and ``mother'')
but are far less common.
Another point is that CVTs use a~lot of facts to tell the same thing that
would be represented by a~single Wikidata statement.
For example, to encode that Barack Obama is the president of the United States from June~20, 2009,
Freebase requires more than six facts as shown in \autoref{fig:cvt-obama}.
If we try to encode Wikidata statements as if they were Freebase facts, \emph{i.e.},
by removing sources, representing statements with qualifiers using CVTs
and adding reverse properties it would lead to 110M of facts,
\emph{i.e.}, 167\% of the raw number of statements.
Sometimes, Freebase contains also duplicate data.
For example, a~lot of cities in the United States have duplicate CVTs for the population
where the only change is the encoding of the source
%\footnote{See, for example, \url{http://www.freebase.com/m/01jr6}.}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/wikidata-time-space.png}
\caption{Time and spatial distribution of Wikidata items.
Longitude is in $x$ axis and time, following a~power law ($y' = y^8$), is in $y$ axis}
\label{fig:time-space-wikidata}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/freebase-time-space.png}
\caption{Same figure for Freebase entities}
\label{fig:time-space-freebase}
\end{figure}

We have also created a~time-spatial visualization of Freebase and Wikidata in order
to get an idea of the coverage of the two knowledge bases
(see \autoref{fig:time-space-wikidata} and \autoref{fig:time-space-freebase}).
The used method is to extract the earliest date and the longitude
linked to each knowledge base entities
and propagate them to the connected entities which do not have such data
using a~decreasing exponential distribution.
We see that the two knowledge bases follow the same basic patterns,
with a~coverage mostly concentrated on Europe and North America,
and there are no strong differences between them.
The coverage of the after 2000 entities seems better in Freebase, maybe because of
the presence of a~lot of musical data missing in Wikidata and of the U.S.
population CVTs.
The big horizontal lines seems to be mostly due on places that do not have any linked date
and so inherit the foundation date of their country.
For Wikidata, two lines that are caused by this behavior are the big line over the U.S.A around 1780
and the one covering Russia around 850.
We could conclude from this visualization that Wikidata, even with its smaller size,
seems to have a~nearly as good coverage of the world as Freebase.
To conclude, it is tricky to find good ways to compare the two databases
and good metrics of success for the migration and we could state that, for a~knowledge base,
containing far more triples than the other does not mean
it contains the same proportion of additional good data.

>>>>

\usepackage[utf8]{inputenc}
\usepackage[hyphens]{url}
\usepackage[pdftex,urlcolor=black,colorlinks=true,linkcolor=black,citecolor=black]{hyperref}
\def\sectionautorefname{Section}
\def\subsectionautorefname{Subsection}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{gensymb}
% todo macro
\usepackage{color}
\newcommand{\todo}[1]{\noindent\textcolor{red}{{\bf \{TODO}: #1{\bf \}}}}

@@

\subsection{Quantitative Comparison}

In order to put the statistics in this section in context,
in the following we provide an overview of the size of the last dump of Freebase
from March 2015:

\begin{itemize}
  \setlength\itemsep{0em}
  \item 48M~topics
  \item 2,997M triples
  \item 442M facts owned by Google%
    \footnote{These facts are the remaining Freebase triples
      after filtering out data about IDs (\texttt{/type/object/id}),
      types (\texttt{/type/object/type}), labels, descriptions (\texttt{/type/object/name},
      \texttt{/common/topic/description}, \emph{etc.}), and triples that were removed
      due to their licensing (see \autoref{sec:licensing})}
  \item 68M~labels
\end{itemize}

As a~direct comparison, below the raw statistics for Wikidata as of August 2015
that make the differences in pure size between the two knowledge bases quite evident.

@@

Why are these numbers so different?
Regarding the number of topics, Freebase has a~lot of topics about subjects
that do not match Wikidata's notability criteria.%
\footnote{Wikidata notability criteria:
\url{https://www.wikidata.org/wiki/Wikidata:Notability}}
For example, Freebase holds data about 11.3M~musical recordings,
8.3M~music canonical versions, 2.6M~ISBNs, \emph{etc.} that are not contained in Wikidata.
We also cannot compare the number of facts and the number of statements directly,
even considering the lower number of topics.
Freebase encodes its knowledge far more redundantly than Wikidata.
To illustrate this, properties in Freebase often have a~\emph{reverse} property
that is used in order to be able to traverse the Freebase graph easily in both directions.
For example, the property \texttt{/people/person/place\_of\_birth} has the corresponding
reverse property \texttt{/location/location/people\_born\_here}
that encodes exactly the same semantic information.
Such reverse properties sometimes exist in Wikidata
(like \texttt{children} for \texttt{father} and \texttt{mother}),
but are far less common.
Another point is that CVTs use a~lot of facts to convey the exact same information
that can be represented with a~single Wikidata statement.
For example, to encode that Barack Obama is the president of the United States as of June~20, 2009,
Freebase requires more than six facts as shown in \autoref{fig:cvt-obama}.
If we attempted to encode Wikidata statements as if they were Freebase facts, \emph{i.e.},
by removing sources, representing statements with qualifiers using CVTs,
and adding reverse properties, this would lead to 110M facts,
\emph{i.e.}, 167\% of the raw number of statements.
In a~considerable amount of cases, Freebase also contains duplicate data.
For example, many cities in the United States have duplicate CVTs for the population,
where the only change is the encoding of the source.

\subsection{Spatio-Temporal Comparison}

In order to get an idea of the coverage of the two knowledge bases,
we have further created a~spatio-temporal visualization of Freebase and Wikidata 
that is shown in \autoref{fig:time-space}.
The used method is to extract the earliest date and the longitude
linked to each of the entities of the two knowledge bases
and propagate them to the connected entities which do not have such data
using a~decreasing exponential distribution \todo{Please clarify the ``propagate'' part}.
We see that the two knowledge bases follow the same basic patterns,
with their coverage mostly concentrated on Europe and North America.
There are no strong differences between them,
yet coverage of \mbox{\emph{post}-year-2000} entities seems better in Freebase.
Our theory is that this is \emph{(i)}~due to the absence of a~lot of musical data in Wikidata
that in contrast is present in Freebase,
and \emph{(ii)} due to the US population CVTs in Freebase.
The big horizontal lines in \autoref{fig:freebase} and more even in \autoref{fig:wikidata}
can most probably be traced back to places that do not have any linked date,
and in consequence inherit the foundation date of their country.
For Wikidata (\autoref{fig:wikidata}), two lines that are definitely caused by this effect
are the big line over the USA around 1780
and the one covering Russia around 850.
One possible conclusion from this visualization could be that Wikidata,
despite its smaller absolute size, appears to have
a~nearly as complete coverage of the world as Freebase.
Nonetheless, it is very challenging to find objective ways to compare the two databases
and good metrics of success for the migration.
We certainly note that for knowledge bases the amount of triples is not an adequate measure
to judge the proportion of additional good data.

\begin{figure*}[!htbp]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{img/freebase-time-space.png}
        \caption{Freebase}
        \label{fig:freebase}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{img/wikidata-time-space.png}
        \caption{Wikidata}
        \label{fig:wikidata}
    \end{subfigure}
    \caption{Spatio-temporal distribution of Freebase and Wikidata items,
      the $x$ axis shows longitude degrees from 180°~E to 180°~W, the $y$ axis shows time from the year 0 to 2015, following a~power law ($y' = y^8$)}
    \label{fig:time-space}
\end{figure*}
